\chapter{Software}

\section{Preparation}
\label{preparation}

Do not underestimate the effort required to flash your first LED. You
require:

\begin{itemize}
\item
  A computer with \url{git} installed and a useful shell program such as
  \url{bash}. UC has a \href{http://ucmirror.canterbury.ac.nz/}{mirror}
  for a variety of Linux distributions; we recommend Ubuntu or Mint.
\item
  A working \href{ARM_toolchain}{ARM toolchain} (arm-none-eabi-gcc/g++
  version 4.9.3 or newer).
\item
  \url{OpenOCD} V0.8.0 or later.
\item
  An \href{ST-link_programmer}{ST-link programmer} and 10-wire ribbon
  cable for programming. You can get the adaptor from Scott Lloyd in the
  SMT lab. You will need to make your own cable (For a grey ribbon
  cable, align the red stripe with the small arrow denoting pin 1 on the
  connector. For a rainbow ribbon cable, connect the brown wire to pin
  1.). There are two variants of the \href{ST-link_programmer}{ST-link
  programmer} with \textbf{different pinouts} so you may need to
  customise your programming cable.
\item
  Plenty of gumption.
\end{itemize}

\subsection{Toolchain}
\label{toolchain}

The required software is installed on computers in the \url{ESL} and
\url{CAE}. It should run under both Linux and Windows. If there is a
problem ask the technical staff.

If you are working from a personal Linux computer make sure to update
your machine to the latest software versions. This is particularly
important for the ARM toolchain (arm-none-eabi-gcc/g++) which should be
at version 4.9.3 or newer. You can do this with the following command
for an Ubuntu or Mint distribution:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{sudo}\NormalTok{ apt update }\KeywordTok{&&} \FunctionTok{sudo}\NormalTok{ apt upgrade}
\end{Highlighting}
\end{Shaded}

For MacOS machines that have \href{https://brew.sh}{homebrew} installed,
you can use the following command:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{brew}\NormalTok{ install openocd git}
\ExtensionTok{brew}\NormalTok{ cask install gcc-arm-embedded}
\end{Highlighting}
\end{Shaded}

\subsection{Project forking}
\label{project-forking}

The example software is hosted on the eng-git \url{Git} server at
\href{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021}{\url{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021}}.
Your group leader should have forked this template project. This creates
your own group copy of the project on the eng-git server that you can
modify, add members, etc. To fork the project:

\begin{enumerate}
\item
  Go to
  \href{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021}{\url{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021}}.
\item
  Click the `Fork' button. This will create a copy of the main repository
  for the project.
\item
  Click on the `Settings' menu then click the `Expand' button for
  `Sharing and permissions'. Change `Project Visibility' to `Private'.
\item
  Click on the `Members' menu and add group members as Developers.
\end{enumerate}

\subsection{Project cloning}
\label{project-cloning}

Once your project has been forked from the template project, each group
member needs to clone it. This makes a local copy of your project on
your computer. \textbf{If inside the university, you will need to
\href{UC_internet_enabler}{enable the internet}} since two
\href{git_submodules}{git submodules} are required.

If you are using an ECE computer, it is advised that you clone the
project on to a removable USB flash drive. This will make git operations
and compilation a 100 times faster than using the networked file system.

There are two ways to clone the project. If you are impatient and do not
mind having to enter a username and password for every git pull and push
operation use:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{git}\NormalTok{ clone https://eng-git.canterbury.ac.nz/groupleader-userid/wacky-racers-2021.git wacky-racers}
\end{Highlighting}
\end{Shaded}

Otherwise, set up \url{ssh-keys} and use:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{git}\NormalTok{ clone git@eng-git.canterbury.ac.nz:groupleader-userid/wacky-racers-2021.git wacky-racers}
\end{Highlighting}
\end{Shaded}

Note, the recursive option should clone the submodules \url{mmculib} and
\url{mat91lib} from github.com.

You can several different cloned copies of your project in different
directories. Sometimes if you feel that the world, and \url{git} in
particular, is against you, clone a new copy, using:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{git}\NormalTok{ clone https://eng-git.canterbury.ac.nz/groupleader-userid/wacky-racers-2021.git wacky-racers-new}
\end{Highlighting}
\end{Shaded}

\subsection{Configuration}
\label{configuration}

Each board has different PIO definitions and requires its own
configuration information. The \texttt{wacky-racers/src/boards}
directory contains a number of configurations; one for the hat and racer
boards. Each configuration directory contains three files:

\begin{itemize}
\item
  \texttt{board.mk} is a makefile fragment that specifies the MCU model,
  optimisation level, etc.
\item
  \texttt{target.h} is a C header file that defines the PIO pins and
  clock speeds.
\item
  \texttt{config.h} is a C header file that wraps target.h. It's purpose
  is for porting to different compilers.
\end{itemize}

\textbf{You will need to edit the target.h file for your board} and set
the definitions appropriate for your hardware. Here's an excerpt from
\texttt{target.h}:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{/* USB  */}
\PreprocessorTok{#define USB_VBUS_PIO PA5_PIO}

\CommentTok{/* ADC  */}
\PreprocessorTok{#define ADC_BATTERY PB3_PIO}
\PreprocessorTok{#define ADC_JOYSTICK_X PB2_PIO}
\PreprocessorTok{#define ADC_JOYSTICK_Y PB1_PIO}

\CommentTok{/* IMU  */}
\PreprocessorTok{#define IMU_INT_PIO PA0_PIO}

\CommentTok{/* LEDs  */}
\PreprocessorTok{#define LED1_PIO PA20_PIO}
\PreprocessorTok{#define LED2_PIO PA23_PIO}
\end{Highlighting}
\end{Shaded}

\section{First program}
\label{first-program}

Your first program to test your board should only flash an LED (the
hello world equivalent for embedded systems). The key to testing new
hardware is to have many programs that only do one simple task each.

\subsection{OpenOCD}
\label{openocd}

\url{OpenOCD} is used to program the SAM4S. For this assignment, we are
using a \href{ST-link_programmer}{ST-link programmer} to connect to the
SAM4S using serial wire debug (SWD). This connects to your board with a
10-wire ribbon cable and an \href{IDC_connector}{IDC connector}.

\begin{enumerate}
\item
  Before you start, disconnect the battery and other cables from your
  PCB.
\item
  Connect a 10-wire ribbon cable from the
  \href{ST-link_programmer}{ST-link programmer} to the programming
  header on your PCB. This will provide 3.3 V to your board so your
  green power LED should light.
\item
  Open a \textbf{new terminal window} and start \url{OpenOCD}.
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ wacky-racers}
\NormalTok{$ }\ExtensionTok{openocd}\NormalTok{ -f src/mat91lib/sam4s/scripts/sam4s_stlink.cfg}
\end{Highlighting}
\end{Shaded}

All going well, the last line output from \url{OpenOCD} should be:

\begin{verbatim}
Info : sam4.cpu: hardware has 6 breakpoints, 4 watchpoints
\end{verbatim}

Congrats if you get this! It means you have correctly soldered your
SAM4S. If not, do not despair and do not remove your SAM4S. Instead,
see \protect\hyperref[troubleshooting]{troubleshooting}.


\subsection{LED flash program}
\label{led-flash-program}

For your first program, use
\texttt{wacky-racers/src/demo-apps/led\_test1/led\_test1.c}. The macros
\texttt{LED1\_PIO} and \texttt{LED2\_PIO} need to be defined in
\texttt{target.h} (see
\protect\hyperref[configuration]{configuration}). If your LEDs are
active high, replace \texttt{ActiveLow} with \texttt{ActiveHigh}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{/* File:   led_test1.c}
\CommentTok{   Author: M. P. Hayes, UCECE}
\CommentTok{   Date:   25 Feb 2018}
\CommentTok{   Descr: }
\CommentTok{*/}
\PreprocessorTok{#include }\ImportTok{"led.h"}
\PreprocessorTok{#include }\ImportTok{"pacer.h"}

\PreprocessorTok{#define PACER_RATE 10}

\DataTypeTok{int}\NormalTok{ main (}\DataTypeTok{void}\NormalTok{)}
\NormalTok{\{}
\NormalTok{    led_cfg_t led1 = \{LED1_PIO, }\DecValTok{0}\NormalTok{\};}
\NormalTok{    led_cfg_t led2 = \{LED2_PIO, }\DecValTok{0}\NormalTok{\};}

\NormalTok{    pacer_init(PACER_RATE);}
\NormalTok{    led_init(&led1);}
\NormalTok{    led_init(&led2);}

\NormalTok{    led_set (&led1, }\DecValTok{1}\NormalTok{);}

    \ControlFlowTok{while}\NormalTok{ (}\DecValTok{1}\NormalTok{)}
\NormalTok{    \{}
\NormalTok{        pacer_wait ();}
\NormalTok{        led_toggle (&led2);}
\NormalTok{    \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\subsection{Compilation}
\label{compilation}

Due to the many files required, compilation is performed using
makefiles.

The demo test programs are generic and you need to specify which board
you are compiling them for. The board configuration file can be chosen
dynamically by defining the environment variable \texttt{BOARD}. For
example,

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ src/test-apps/ledflash1}
\NormalTok{$ }\VariableTok{BOARD=}\NormalTok{racer }\FunctionTok{make}
\end{Highlighting}
\end{Shaded}

If all goes well, you should see at the end:

\begin{verbatim}
   text    data     bss     dec     hex filename
  19432    2420    2732   24584    6008 led_test1.bin
\end{verbatim}

To avoid having to specify the environment variable \texttt{BOARD}, you
can define it for the rest of your session using:
%
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\BuiltInTok{export} \VariableTok{BOARD=}\NormalTok{racer}
\end{Highlighting}
\end{Shaded}
%
and then just use
%
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{make}
\end{Highlighting}
\end{Shaded}

Note, if you compile with the \texttt{BOARD} environment variable set incorrectly, use
\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{make clean}
\NormalTok{$ }\FunctionTok{make}  
\end{Highlighting}
\end{Shaded}
%
to delete all the object files and rebuild.


\subsection{Booting from flash memory}
\label{booting-from-flash-memory}

By default the SAM4S runs a bootloader program stored in ROM. The SAM4S
needs to be configured to run your application in flash memory.

If \url{OpenOCD} is running you can do this with:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{make}\NormalTok{ bootflash}
\end{Highlighting}
\end{Shaded}

Unless you force a complete erasure of the SAM4S flash memory by
connecting the \texttt{ERASE} pin to 3.3 V, you will not need to repeat
this command.

\subsection{Programming}
\label{programming}

If \url{OpenOCD} is running you can store your program in the flash
memory of the \url{SAM4S} using:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{make}\NormalTok{ program}
\end{Highlighting}
\end{Shaded}

When this finishes, one of your LEDs should flash. If so, congrats! If
not, see \protect\hyperref[troubleshooting]{troubleshooting}.

To reset your SAM4S, you can use:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{make}\NormalTok{ reset}
\end{Highlighting}
\end{Shaded}

\section{USB interfacing}
\label{usb-interfacing}

To help debug your programs, it is is wise to set up \href{USB_CDC}{USB
CDC}. For example, here's the code for
\texttt{wacky-racers/src/demo-apps/usbserial\_hello1/usbserial\_hello1.c}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{/* File:   usbserial_hello1.c}
\CommentTok{   Author: M. P. Hayes, UCECE}
\CommentTok{   Date:   22 Feb 2018}
\CommentTok{*/}
\PreprocessorTok{#include }\ImportTok{<stdio.h>}
\PreprocessorTok{#include }\ImportTok{"usb_serial.h"}
\PreprocessorTok{#include }\ImportTok{"delay.h"}

\DataTypeTok{static}\NormalTok{ usb_serial_cfg_t usb_serial_cfg =}
\NormalTok{\{}
\NormalTok{    .read_timeout_us = }\DecValTok{1}\NormalTok{,}
\NormalTok{    .write_timeout_us = }\DecValTok{1}\NormalTok{,}
\NormalTok{\};}

\DataTypeTok{int}\NormalTok{ main (}\DataTypeTok{void}\NormalTok{)}
\NormalTok{\{}
    \DataTypeTok{int}\NormalTok{ i;}

    \CommentTok{// Create non-blocking tty device for USB CDC connection.}
\NormalTok{    usb_serial_init (&usb_serial_cfg, }\StringTok{"/dev/usb_tty"}\NormalTok{);}
    
    \CommentTok{// Redirect printf to USB serial}
\NormalTok{    freopen (}\StringTok{"/dev/usb_tty"}\NormalTok{, }\StringTok{"a"}\NormalTok{, stdout);}
\NormalTok{    freopen (}\StringTok{"/dev/usb_tty"}\NormalTok{, }\StringTok{"r"}\NormalTok{, stdin);    }

    \ControlFlowTok{for}\NormalTok{ (i = }\DecValTok{0}\NormalTok{; i < }\DecValTok{100}\NormalTok{; i++)}
\NormalTok{    \{}
\NormalTok{        printf (}\StringTok{"Hello world %d}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, i);}
\NormalTok{        fflush (stdout);}
\NormalTok{        delay_ms (}\DecValTok{100}\NormalTok{);}
\NormalTok{    \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

To get this to work you need to compile it and program the SAM4S using:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ wacky-racers/src/demo-apps/usbserial_hello1}
\NormalTok{$ }\FunctionTok{make}\NormalTok{ program}
\end{Highlighting}
\end{Shaded}

You then need to connect your computer to the USB connector on your PCB.
If you are running Linux, run

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{dmesg}
\end{Highlighting}
\end{Shaded}

This should say something like:

\begin{verbatim}
Apr 30 11:03:50 thing4 kernel: [52704.481352] usb 2-3.3: New USB device found, idVendor=03eb, idProduct=6202
Apr 30 11:03:50 thing4 kernel: [52704.481357] usb 2-3.3: New USB device strings: Mfr=1, Product=2, SerialNumber=3
Apr 30 11:03:50 thing4 kernel: [52704.482060] cdc_acm 2-3.3:1.0: ttyACM0: USB ACM device
\end{verbatim}

Ignore the errors for now (they have popped up with new Linux kernels
and I have yet to determine why). If you see
\texttt{ttyACM0:\ USB\ ACM\ device} congrats. If not, see
\href{USB_debugging}{USB debugging}.

You can now run a \href{Serial_terminal_applications}{serial terminal
program}. For example,

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{gtkterm}\NormalTok{ -p /dev/ttyACM0}
\end{Highlighting}
\end{Shaded}

All going well, this will repeatedly print 'Hello world'.

If you get an error 'device is busy', it is likely that the ModemManager
program has automatically connected to your device on the sly. This
program should be disabled on the ECE computers. For more about this and
using other operating systems, search for USB CDC on ecewiki.

\section{Test programs}
\label{test-programs}

There are a number of test programs in the directory
\texttt{wacky-racers/demo-apps}. Where possible these are written to be
independent of the target board using configuration files (see
\protect\hyperref[configuration]{configuration}).

\subsection{PWM test}
\label{pwm-test}

The program
\href{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021/blob/master/src/test-apps/pwm_test2/pwm_test2.c}{\texttt{src/test-apps/pwm\_test2/pwm\_test2.c}}
provides an example of driving PWM signals. Notes:
%
\begin{enumerate}
\item
  This is for a different H-bridge module that requires two PWM signals
  and forward/reverse signals. You will need to generate four PWM
  signals or be clever with two PWM signals.
\item
  The \texttt{pwm\_cfg\_t} structure configures the frequency, duty
  cycle and alignment of the output PWM.
\end{enumerate}

The most likely problem is that you have not used a PIO pin that can be
driven as a PWM output. The SAM4S can generate four independent hardware
PWM signals. See
\href{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021/blob/master/src/mat91lib/pwm/pwm.c}{\texttt{src/mat91lib/pwm/pwm.c}}
for a list of supported PIO pins. Note, PA16, PA30, and PB13 are
different options for PWM2.

To drive the motors you will need to use a
\href{bench_power_supply}{bench power supply}. Start with the current
limit set at 100\,mA maximum in case there are any board shorts.
When all is well, you can increase the current limit.

\subsection{IMU test}
\label{imu-test}

The program
\href{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021/blob/master/src/demo-apps/imu_test1/imu_test1.c}{\texttt{wacky-racers/src/demo-apps/imu\_test1/imu\_test1.c}}
provides an example of using the MPU9250 IMU. All going well, this
prints three 16-bit acceleration values per line to USB CDC. Tip your
board over, and the the third (z-axis) value should go negative since
this measures the effect of gravity on a little mass inside the IMU
pulling on a spring.

If you get 'Cannot detect IMU' the main reasons are:

\begin{enumerate}
\item
  You have specified the incorrect address. Use \texttt{MPU\_0} if the
  AD0 pin is connected to ground otherwise use \texttt{MPU\_1}.
\item
  You are using TWI1. The PB4 and PB5 pins used by TWI1 default to JTAG
  pins. See \protect\hyperref[disabling-jtag-pins]{disabling JTAG pins}.
\item
  You do not have TWI/I2C pull-up resistors connected to 3V3 or the
  wrong resistor values. Use a scope to look at the signals TWCK/SCLK
  and TWD/SDA.
\item
  Check that the auxiliary \url{I2C} bus signals are not connected
  (otherwise the magnetometer will not respond).  
\item
  You have a soldering problem. Try pressing on a side of the IMU with a
  fingernail to see if it starts working.
\end{enumerate}

If you reset the \url{SAM4S} in the middle of a transaction with the
IMU, the IMU gets confused and holds the SDA/TWD line low. This requires
recycling of the power or sending out some dummy clocks on the SCL/TWCK
signal.


\subsection{Radio test}
\label{radio-test}

The program
\href{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021/blob/master/src/demo-apps/radio_tx_test1/radio_tx_test1.c}{\texttt{wacky-racers/src/demo-apps/radio\_tx\_test1/radio\_tx\_test1.c}}
provides an example of using the radio as a transmitter.

The companion program
\href{https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021/blob/master/src/demo-apps/radio_rx_test1/radio_rx_test1.c}{\texttt{wacky-racers/src/demo-apps/radio\_rx\_test1/radio\_rx\_test1.c}}
provides an example of using the radio as a receiver.

Notes:
%
\begin{enumerate}
\item
  Both programs must use the same RF channel and the same address. Some
  RF channels are better than others since some overlap with \url{WiFi}
  and \url{Bluetooth}. The address is used to distinguish devices
  operating on the same channel. Note, the transmitter expects an
  acknowledge from a receiver on the same address and channel.
\item
  The radio `write` method blocks waiting for an auto-acknowledgement
  from the receiver device. This acknowledgement is performed in
  hardware. If no acknowledgement is received, it retries for up to 15
  times. The auto-acknowledgement and number of retries can be
  configured in software.
\end{enumerate}

If you cannot communicate between your hat and racer boards, try
communicating with the radio test modules Scott Lloyd has in the SMT
lab.

\section{Your hat/racer program}
\label{your-hatracer-program}

There is skeleton code, see
\texttt{wacky-racers/src/wacky-apps/hat/hat.c} and
\texttt{wacky-racers/src/wacky-apps/racer/racer.c}. We recommend that
build your programs incrementally and that you poll your devices with a
paced loop and not use interrupts. It is a good idea to disable the
watchdog timer until you have robust code.

\begin{Shaded}
\begin{Highlighting}[]
\PreprocessorTok{#include }\ImportTok{<stdio.h>}
\PreprocessorTok{#include }\ImportTok{<math.h>}
\PreprocessorTok{#include }\ImportTok{"mcu.h"}
\PreprocessorTok{#include }\ImportTok{"pacer.h"}

\CommentTok{// Set to the desired polling rate.}
\PreprocessorTok{#define PACER_RATE 1000}

\DataTypeTok{int}
\NormalTok{main (}\DataTypeTok{void}\NormalTok{)}
\NormalTok{\{}
    \CommentTok{// If you are using PB4 or PB5 (say for TWI1) uncomment the next line.}
    \CommentTok{// mcu_jtag_disable ();    }

\NormalTok{    pacer_init (PACER_RATE);    }
    
\NormalTok{    mcu_watchdog_enable ();}

    \ControlFlowTok{while}\NormalTok{ (}\DecValTok{1}\NormalTok{)}
\NormalTok{    \{}
\NormalTok{        pacer_wait ();}

\NormalTok{        mcu_watchdog_reset ();}

        \CommentTok{/* Do your stuff here...  */}
        
\NormalTok{    \}}
    
    \ControlFlowTok{return} \DecValTok{0}\NormalTok{;}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\section{Debugging}
\label{debugging}

If \url{OpenOCD} is running, a running program can be debugged using:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{make}\NormalTok{ debug}
\end{Highlighting}
\end{Shaded}

This starts \url{GDB} and attaches to the target MCU. \url{GDB} is a
command line debugger but there are many GUI programs that will control
it, for example, \url{vscode} and \url{geany} have plugins.

\url{GDB} allows you to inspect the CPU registers, memory, set
breakpoints, set watchpoints, and much more. The only gnarly thing is
that when compiling with optimisation, the compiler may reorder
statements or even remove them. Debugging embedded systems is also made
difficult by the asynchronous nature of I/O.

You can reset your program using the GDB `jump reset` command. However,
this does not reset the peripherals as with a power-on reset.

\section{Howtos}
\label{howtos}

\subsection{Disabling JTAG pins}
\label{disabling-jtag-pins}

By default PB4 and PB5 are configured as \url{JTAG} pins. You can turn
them into PIO pins or use them for TWI1 using:

\begin{Shaded}
\begin{Highlighting}[]
\PreprocessorTok{#include }\ImportTok{"mcu.h"}

\DataTypeTok{void}\NormalTok{ main (}\DataTypeTok{void}\NormalTok{)}
\NormalTok{\{}
\NormalTok{    mcu_jtag_disable ();}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\subsection{Watchdog timer}
\label{watchdog-timer}

The watchdog timer is useful for resetting the SAM4S if it hangs in a
loop.  It is disabled by default but can be enabled using:

\begin{Shaded}
\begin{Highlighting}[]
\PreprocessorTok{#include }\ImportTok{"mcu.h"}

\DataTypeTok{void}\NormalTok{ main (}\DataTypeTok{void}\NormalTok{)}
\NormalTok{\{}
\NormalTok{    mcu_watchdog_enable ();}
   
    \ControlFlowTok{while}\NormalTok{ (}\DecValTok{1}\NormalTok{)}
\NormalTok{    \{}
         \CommentTok{/* Do your stuff here.  */}

\NormalTok{         mcu_watchdog_reset ();}
\NormalTok{    \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\section{Under the bonnet}
\label{under-the-bonnet}

wackylib is a library of C files that tries to hide some of the
complexity of the underlying software.

mmculib is a library of C drivers, mostly for performing high-level I/O.
It is written to be microcontroller neutral.

mat91lib is a library of C drivers specifically for interfacing with the
peripherals of Atmel AT91 microcontrollers such as the
\href{Atmel_SAM4S}{Atmel SAM4S}. It provides the hardware abstraction
layer.

The building is controlled by \texttt{mat91lib/mat91lib.mk}. This is a
makefile fragment loaded by \texttt{mmculib/mmculib.mk}.
\texttt{mat91lib/mat91lib.mk} loads other makefile fragments for each
peripheral or driver required. It also automatically generates
dependency files for the gazillions of other files that are required to
make things work.

\section{Git}
\label{git}

To properly use git you should commit and push often. The smaller the
changes and the more often you make per commit, the smaller the chance
of the dreaded merge conflict.

\subsection{Git pulling from upstream}
\label{git-pulling-from-upstream}

To be able to get updates if the template project is modified you will
need to:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ wacky-racers }
\NormalTok{$ }\FunctionTok{git}\NormalTok{ remote add upstream https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021.git  }
\end{Highlighting}
\end{Shaded}

Again if you do not want to manually enter your password (and have
ssh-keys uploaded) you can use:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ wacky-racers }
\NormalTok{$ }\FunctionTok{git}\NormalTok{ remote add upstream git@eng-git.canterbury.ac.nz:mph/wacky-racers-2021.git    }
\end{Highlighting}
\end{Shaded}

Once you have defined the upstream source, to get the updates from the
main repository use:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{git}\NormalTok{ pull upstream master}
\end{Highlighting}
\end{Shaded}

If you enter the wrong URL make a mistake, you can list the remote
servers and delete the dodgy entry using:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{git}\NormalTok{ remote -v}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ remote rm upstream}
\end{Highlighting}
\end{Shaded}

Note, \textbf{origin} refers to your group project and \textbf{upstream}
refers to the template project that origin was forked from.

\begin{itemize}
\item
  \emph{\url{git} does not update to the latest version of the mat91lib
  and mmculib submodules}. We are not sure why. As a hack use:
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ wacky-racers}
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ src/mat91lib}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ checkout master}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ pull origin master}
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ ../mmculib}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ checkout master}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ pull origin master}
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ ..}
\NormalTok{$ }\BuiltInTok{cd}\NormalTok{ demo-apps}
\NormalTok{$ }\FunctionTok{make}\NormalTok{ clean}
\end{Highlighting}
\end{Shaded}

\subsection{Git merging}
\label{git-merging}

The bane of all version control programs is dealing with a merge
conflict. You can reduce the chance of this happening by committing and
pushing faster than other people in your group.

If you get a message such as:

\begin{verbatim}
From https://eng-git.canterbury.ac.nz/mph/wacky-racers-2021
 * branch            master     -> FETCH_HEAD
error: Your local changes to the following files would be overwritten by merge:
    src/demo-apps/imu_test1/imu_test1.c
Please, commit your changes or stash them before you can merge.
\end{verbatim}

what you should so is:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{$ }\FunctionTok{git}\NormalTok{ stash}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ pull}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ stash pop}
\CommentTok{# You may now have a merge error.  You will now have to edit the offending file, in this case imu_test1.c}
\CommentTok{# Once the file has been fixed}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ add src/demo-apps/imu_test1/imu_test1.c}
\NormalTok{$ }\FunctionTok{git}\NormalTok{ commit -m }\StringTok{"Fix merge"}
\end{Highlighting}
\end{Shaded}

Sometimes when you do a git pull you will be thrown into a text editor
to type a merge comment. The choice of editor is controlled by an
environment variable \texttt{EDITOR}. On the ECE computers this defaults
to emacs. You can changed this by adding a line such as the following to
the \texttt{.bash\_profile} file in your home directory.

\begin{Shaded}
\begin{Highlighting}[]
\BuiltInTok{export} \VariableTok{EDITOR=}\NormalTok{geany}
\end{Highlighting}
\end{Shaded}

By the way, to exit emacs type ctrl-x ctrl-c

\textbf{Please do not edit the files in the mavrlib, mmculib, and
wackylib} directories since this can lead to merge problems in the
future. If you find a bug or would like additional functionality let MPH
or one of the TAs know.
